---
title: "Cross-situational word learning with Brazilians"
output: html_notebook
author: "Rodrigo Dal Ben"
date: 10/02/2021
---

Last update: 2021-11-15

# Background

#ADD DETAILS
**Analyses reported by**: Dal Ben, R., Souza, D. H. & Hay, J. (under review). Combining statistics: The role of phonotactics on cross-situational word learning. ##### ADD DETAILS HERE ######

**OSF project**: https://osf.io/6fqzg/

**Feedback and suggestions:** <dalbenwork@gmail.com>

# Last packages update

```{r}
# YYYY-MM-DD
last_upd <- "2021-11-10" 
```

# Load packages

Here we use `groundhog` package to minimize conflicts among packages versions.

```{r}
# load 'groundhog', version:1.5.0, using R-4.1.2
library(groundhog)

# set vector w/ packages
packages = c("tidyverse",
             "tidylog", 
             "boot",
             "gmodels", 
             "lme4", 
             "brms",
             "sjPlot", 
             "patchwork", 
             "janitor",
             "RColorBrewer", 
             "beepr", 
             "here")

# load packages
groundhog::groundhog.library(pkg = packages, date = last_upd)

# avoid scientific notation
options(scipen = 0)

# color-blind palette
cb_pal = c("#E69F00", "#009E73", "#0072B2", "#D55E00", "#999999", "#000000") 
```

# Load data

```{r}
# load 
load(here("02_data/tidy_data_cswl_br.rda"))

# double check sample size (should be 30)
data %>% distinct(p_n_unique) %>% count() # 30 participants
```

# Prepare data

```{r}
# filter test trials and set index for trials with rt > 3 SDs
sd_comp <- 
  data %>% 
  filter(trial_type == "test cswl") %>% 
  summarise(sd_1 = sd(rt, na.rm = T),
            sd_3 = sd_1*3)

data_clean_test <- 
  data %>%
  filter(trial_type == "test cswl") %>% 
  filter(rt < mean(rt, na.rm = T) + sd_comp$sd_3) # 16 trials (2%) excluded

## excluded trials summary
data_clean_test %>% distinct(p_n_unique) %>% count() # 30 participants
data_clean_test %>% group_by(p_n_unique) %>% summarise(n_trials = n())
```

# Demographics

```{r}
# summary data
data_clean_test %>% 
  group_by(p_n_unique) %>% 
  slice_head() %>%
  ungroup() %>% 
  summarise(
    n = n(), 
    age_avg = mean(age),
    age_sd = sd(age),
    female = length(which(gender == "female")),
  )
```

# Visualizations

## Summary data

```{r}
# calculating mean for each block
data_summ <- 
  data_clean_test %>% 
  group_by(p_n_unique, stimuli_set_pp) %>% 
  summarise(
    is_correct_avg = mean(is_correct),
    is_correct_sum = sum(is_correct),
    trials_n = 24,
    rt = mean(rt)
    ) %>% 
  ungroup()

# number of trials and proportion
data_summ %>% group_by(stimuli_set_pp) %>% summarise(m = mean(is_correct_sum), sd = sd(is_correct_sum))
data_summ %>% group_by(stimuli_set_pp) %>% summarise(m = mean(is_correct_avg), sd = sd(is_correct_avg))
```

## Mean correct 

```{r}
# sum
sum_pp_plot <-  
  data_summ %>% 
  ggplot(aes(x = as_factor(""), y = is_correct_sum, color = stimuli_set_pp)) +
  geom_violin(alpha = 0.6) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.9), 
             alpha = 0.3) + 
  stat_summary(fun.data = "mean_cl_boot", 
               position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(0, 24, 4)) +
  geom_hline(yintercept = 6, linetype = "dashed") +
  labs(y = "Mean number of correct selections", x = "Phonotactics") +
  theme_bw()
        
sum_pp_plot

# proportions
avg_pp_plot <-  
  data_summ %>% 
  ggplot(aes(x = stimuli_set_pp, y = is_correct_avg, fill = stimuli_set_pp)) +
  geom_violin(alpha = 0.3) +
  geom_dotplot(binaxis = "y", stackdir = "up", dotsize = 0.5, alpha = 0.6) + 
  stat_summary(fun.data = "mean_cl_boot", size = 0.5, position = position_nudge(x = -0.05, y = 0)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  scale_fill_manual(values = cb_pal) +
  geom_hline(yintercept = 0.25, linetype = "dashed") +
  labs(y = "Mean number of correct selections", x = "Phonotactics") +
  theme_classic() +
  theme(legend.position = "")
        
avg_pp_plot

#ggsave(here("03_output_graphics/plot_pp.tiff"), 
#       width = 15, height = 8, units = "cm", dpi = 300)

# extract the stat_summary from the graph
plot_stats <- ggplot_build(sum_pp_plot)$`data`[[2]] # trials
plot_stats <- ggplot_build(avg_pp_plot)$`data`[[2]] # proportions
```

# Statistics

## Descriptive

```{r}
stats_pp_summ <- 
  data_summ %>%
  group_by(stimuli_set_pp) %>% 
  summarise(
    is_corr = mean(is_correct_avg),
    sd = sd(is_correct_avg), 
    ci_low = gmodels::ci(is_correct_avg)[2],
    ci_high = gmodels::ci(is_correct_avg)[3],
    rt_mean = mean(rt)
  )

stats_pp_summ
```

## Inferential

### Frequentist 

```{r}
## random slope and intercept - FAILED TO CONVERGE
#mod0_cond_freq_exp1 <- 
#  data_clean_test_exp1 %>% 
#  glmer(is_correct ~ offset(logit(chance_level)) + stimuli_set_pp +
#                (target_visual|p_n_unique),  
#              data = .,
#              family = "binomial")

## random intercept
mod1_pp_freq <- 
  data_clean_test %>% 
  #mutate(stimuli_set_pp = factor(stimuli_set_pp, levels = c("PP+", "PP-"))) %>% # PP+ as baseline
  glmer(is_correct ~ offset(logit(chance_level)) + stimuli_set_pp +
                (1|target_visual) + (1|p_n_unique), 
              data = .,
              family = "binomial")

summary(mod1_pp_freq)
confint(mod1_pp_freq, method = "Wald")
tab_model(mod1_pp_freq, title = "Fixed and random effects for Frequentist analysis [selection ~ chance level + phonotactics + (1|stimuli) + (1|participant)]", show.re.var = T, show.icc = T, show.p = F)#, file = here("03_output_graphics/mod_pp_freq.doc"))
```

### Bayesian

```{r}
mod0_pp_bay <- 
  data_clean_test %>% 
  #mutate(stimuli_set_pp = factor(stimuli_set_pp, levels = c("PP+", "PP-"))) %>% # PP+ as baseline
  brm(is_correct ~ offset(logit(chance_level)) + stimuli_set_pp + (target_visual|p_n_unique),
            data = .,
            family = "bernoulli",
            iter = 4000,
            chains = 4
            )

summary(mod0_pp_bay)
tab_model(mod0_pp_bay, title = "Fixed and random effects Bayesian analysis [selection ~ chance level + phonotactics + (stimuli|participant)]", show.re.var = T, show.icc = T)#, file = here("03_output_graphics/mod_pp_bay.doc"))

# let me know when model is done running
beep(sound = 1, expr = NULL)
```

# THE END 
